---
title: "Unifying individual differences in personality, predictability, and plasticity: a practical guide - 'brms' model code"
author: Rose E. O’Dea, Daniel W.A. Noble, Shinichi Nakagawa
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, tidy = TRUE)
options(digits=2)
```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
#install.packages("devtools")
remotes::install_github("rlesur/klippy")
klippy::klippy(tooltip_message = 'Click to Copy Code', tooltip_success = 'Done', position = 'right', color = "red")
```

# R Packages, Data and Model Fits

Readers will  need to download the data and model fits (unless they want to rerun the models) from our [Open Science Framework (OSF) Project Site](http://doi.org/10.17605/OSF.IO/V3QAX). Once downloaded and placed in your working directory run the following code to load the datasets and R packages:

```{r loadpackages, echo = TRUE, results = "hide", messages = FALSE, class.source='klippy'}
# install.packages("pacman")
# Load Libraries and relevant data
		pacman::p_load(brms, tidyverse, coda, MCMCglmm, bayesplot, here, formatR, flextable)	

		
# Get the raw. We need this for back transformation from z-scale
		# making sure fish_ID is coded as a factor (done automatically before R v4+)
		# square-root transforming novel and predator response variables
		# z-scaling all variables
		
	  dat <- read.csv("./Data/data_predictability_worked_example.csv")
		
	  dat <- dat %>% 
	         dplyr::mutate(fish_ID = factor(fish_ID, levels = unique(fish_ID)),
		                       age.Z = scale(age_days),
		                    activity = scale(dist_m_Activity),
		                  aggression = scale(stim.dist_0_5cm_dur_Aggression),
		                       novel = scale(sqrt(stim.dist_0_10cm_dur_Novel)),     
		                    predator = scale(sqrt(stim.dist_0_10cm_dur_Predator)), 
		                      social = scale(stim.dist_0_5cm_dur_Social))	
		
```

# Load Functions 

Load in the functions that will be needed to process output from the models to derive the various estimates needed for calculating the necessary parameters:

```{r funcs, include=TRUE, message=FALSE, warning=FALSE, class.source='klippy'}

  # Population-level intercept   
		func_Bp <- function(B0, B1){
		  Bp <- (2*B0 + B1)/2 # assumes 50/50 ratio for B1 (in our example = sex)
		  return(Bp)
		}

   # Total variance, equation 40. Function for summing dataframe of variance components 
		func_sum_var <- function(vars){
		  sum_var <- sapply(1:nrow(vars), function(x) sum(vars[x,], na.rm = T))
		  return(sum_var)
		}


		# Variance for fixed effects
		func_var_fixed <- function(B1, B2, data){
		  # model matrix for fixed effects
		  X <- model.matrix(~ 1 + sex + age.Z, data = data) 
    		  X1 <- X[,2]
    		  X2 <- X[,3]
		  var_fixed <- sapply(1:length(B1), function(x) var(X1*B1[x] + X2*B2[x]))
		  return(var_fixed)
		}		
		
```
# Running *brms* SHGLM and DHLGM models {.tabset}
## Personality (SHGLM) {.tabset .tabset-fade .tabset-pills} 
### Model {.tabset .tabset-fade .tabset-pills} 
Let's start by fitting a model to the aggression data to estimate "personality".

```{r personality, include=TRUE, message=FALSE, warning=FALSE, class.source='klippy'}
########### Personality ##########
# EXAMPLE 1: standata_personality_shglm_aggression; replicates the following model
#  stan_personality_shglm_aggression <- stan(file = "./Code/stancode_personality_shglm.stan",
#                                          data = standata_personality_shglm_aggression,
#                                          pars = save_pars, 
#                                          iter = 6000, warmup = 2000, chains = 3, cores = 3)
##################################  

	# Fit a model for aggression that includes fixed effects of sex and z-transformed age plus a random intercept for individual ID
		rerun1 = FALSE
		
		if(rerun1) {
		        model1 <- bf(aggression ~ sex + age.Z + (1 | id))
			model1_fit <- brms::brm(model1, data = dat, 
									iter = 6000, warmup = 2000, chains = 3, cores = 3, 
									save_pars = save_pars())
			saveRDS(model1_fit, "./models/model1_fit") 
		} else {
			model1_fit <- readRDS(here("models","model1_fit"))
		}
		
		summary(model1_fit)
```

### Calculate CVm and Rpm {.tabset .tabset-fade .tabset-pills} 

```{r pers_rpm_cv, include=TRUE, message=FALSE, warning=FALSE, class.source='klippy'}

# (Step 1):  Get the average intercept from fixed effects.
Bpm <- func_Bp(    B0 = brms::posterior_samples(model1_fit, pars = "Intercept")[,1], 
		               B1 = brms::posterior_samples(model1_fit, pars = "sex")[,1])

# (Step 2): Extract random intercept for ID. Numerator for eqn 34. Note these are standard deviations for between-individual differences in the mean model
		sigma_m0 <- brms::posterior_samples(model1_fit, pars = "sd_id__Intercept")[,1]
		sigma_m2 <- NA # here we have no random slope, so sigma_m = sigma_m0

# (Step 3): Extract residual standard deviation
		 sigma_e <- brms::posterior_samples(model1_fit, pars = "sigma")[,1]
		
# (Step 4): Calculate fixed effect variance
	var_fixed <- func_var_fixed(  B1 =  brms::posterior_samples(model1_fit, pars = "sex")[,1],
		                            B2 =  brms::posterior_samples(model1_fit, pars = "age.Z")[,1],
		                            data = dat) 

# (Step 5): Calculate total phenotypic variance
		total_var <- func_sum_var(vars = data.frame(sigma_m0^2, var_fixed, sigma_e^2))
		
# (Step 6): Convert back from z-scale 
		
		   mu_unscaled <- mean(dat$stim.dist_0_5cm_dur_Aggression)
		sigma_unscaled <- sd(dat$stim.dist_0_5cm_dur_Aggression)
		
		            Mu <- Bpm*sigma_unscaled + mu_unscaled
	         	 Sigma <- sqrt(total_var)*sigma_unscaled
		
    indiv_sd <- sqrt(sigma_m0^2)*sigma_unscaled

# (Step 7): Calculate CVm 
		CV_m <- indiv_sd/Mu

# (Step 8): Calculate Rpm		
		Rp_m <- sigma_m0^2/total_var

# Results
			personality_results <- data.frame( CVm = mean(CV_m),
			                                   CVm.L = HPDinterval(mcmc(CV_m))[1],
			                                   CVm.U = HPDinterval(mcmc(CV_m))[2],
			                                   	Rpm =  mean(Rp_m),
			                                   Rpm.L = 	HPDinterval(mcmc(Rp_m))[1],
			                                   Rpm.U = 	HPDinterval(mcmc(Rp_m))[1])
			flextable(personality_results)
		
		
```

## Personality / Predictability (DHGLM) {.tabset .tabset-fade .tabset-pills} 
### Model {.tabset .tabset-fade .tabset-pills} 
Let's start by fitting a model to the aggression data to estimate "personality".

```{r personality_pred, include=TRUE, message=FALSE, warning=FALSE, class.source='klippy'}
######################################### Personality / Predictability #####################################
# EXAMPLE 2: standata_personality_dhglm_aggression; replicates the following model
# stan_personality_dhglm_aggression <- stan(file = "./Code/stancode_personality_dhglm.stan",
#                                          data = standata_personality_dhglm_aggression,
#                                          pars = save_pars, 
#                                          iter = 6000, warmup = 2000, chains = 3, cores = 3)	
##########################################################################################################    	

	rerun2=FALSE
	
	if(rerun2){
	    model2 <- bf(aggression ~ sex + age.Z + (1 | p | id),
	    			            sigma ~ sex + age.Z + (1 | p | id))

		model2_fit <- brm(model2, data = dat, 
								iter = 6000, warmup = 2000, chains = 3, cores = 3, 
								save_pars = save_pars()) 
		
		saveRDS(model2_fit, "./models/model2_fit")
	} else {
		model2_fit <- readRDS("./models/model2_fit")
	}

	summary(model2_fit)
```


### Calculate CVm and Rpm {.tabset .tabset-fade .tabset-pills} 